Formulating RDF Constraints Generically and Validating RDF Data Out-Of-The-Box

### 2015-05-08 paper/paper 2/paper 2.tex
feedback Erman

With \emph{minimum qualified cardinality restrictions} (\emph{R-75/81})\footnote{Requirements/constraint types are uniquely identified by alphanumeric technical identifiers like \emph{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}}, researchers from the library domain may restrict
that \emph{publications} must have at least one \emph{author} which must be a \emph{person}.
This constraint can be represented generically (\emph{generic constraint}) in \emph{DL} (\ms{Publication $\sqsubseteq$ $\geq$1 author.Person}), but also specifically (\emph{specific constraint}) by multiple constraint languages:

@@

As every \emph{publication} must have at least one \emph{author} which must be a \emph{person}, 
and since the book \emph{The-Lord-Of-The-Rings} is a \emph{publication} \\(\ms{rdf:type(The-Lord-Of-The-Rings, Publication)}), 
it must have at least one \emph{author} relationship to a \emph{person}.
Constraint violations are raised either (1) if
\emph{The-Lord-Of-The-Rings} does not have any \emph{author} relationship, or (2) if 
it has an \emph{author} which is not a \emph{person} \\
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Hobbit)}), 
or (3) if it has an \emph{author} for which no class is assigned
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}).
In contrast, \emph{The-Lord-Of-The-Rings} is a valid \emph{publication}, if it is connected to a \emph{person} via the property \emph{author}
(\ms{author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Person)}).

@@

%The underlying \emph{semantics} for RDF validation is \emph{UNA} and \emph{CWA}.
RDF validation requires that different names represent different objects (\emph{unique name assumption (UNA)}), whereas,
\emph{OWL 2} is based on the {\em non-unique name assumption (nUNA)}.  
A data property is \emph{functional} (\emph{R-65}) if for each individual \emph{x}, there can be at most one distinct literal \emph{y} such that \emph{x} is connected by the data property to \emph{y}. 
When assuming \emph{UNA}, the functional property {\em title} (\ms{funct (title)}) causes a clash
in case the book {\em Huckleberry-Finn} has more than one \emph{title}.
%(\ms{title(Huckleberry-Finn,} \ms{The-Adventures-of-Huckleberry-Finn)}, \ms{title} \ms{(Huckleberry-Finn,} \ms{Die-Abenteuer-des-Huckleberry-Finn)}).
Assuming \emph{nUNA}, however, reasoning concludes that the titles {\em The-Adventures-of-Huckleberry-Finn} and {\em Die-Abenteuer-des-Huckleberry-Finn} are the same. 
%(\ms{\{The-Adventures-of-Huckleberry-Finn\}} = \ms{\{Die-Abenteuer-des-Huckle} \ms{berry-}\ms{Finn\}}).

Reasoning in \emph{OWL 2} is based on the {\em open-world assumption (OWA)}, i.e., a statement cannot be inferred to be false if it cannot be proved to be true. 
%which fits its primary design purpose to represent knowledge on the World Wide Web. 
%Most of the constraint languages except OWL, have a major shortcoming of redundancy.
%With its open-world semantics, used as a constraint language, OWL solves the redundancy problem. 
%\tb{the following example does not show the redundancy problem!}
%\tb{2. benefit: reasoning solves redundancy problem}
As each \emph{book} must have a \emph{title} (\ms{Book $\sqsubseteq \exists$ title.$\top$}) and {\em Hamlet} is a \emph{book} (\ms{Book(Hamlet)}),
{\em Hamlet} must have at least one \emph{title}.
In an \emph{OWA} setting, this axiom does not cause a constraint violation (even if there is no explicitly defined \emph{title}), since there must be a \emph{title} for this \emph{book} which we may not know. 
On the other hand, RDF validation scenarios require the \emph{closed-world assumption (CWA)}, i.e., a statement is inferred to be false if it cannot be proved to be true.
When assuming the \emph{CWA}, a constraint violation is raised as there is no explicitly defined title. 
A constraint that says that a book must have at least one title in \emph{OWA}, 
does not mean that the title has to be unique. 
In every model  (an interpretation that makes a set of \emph{DL knowledge base} true), 
it can be the case that a book has a single title, however, 
this title is not unique. 
In some models it is \emph{The-Adventures-of-Huckleberry-Finn}, in others it is \emph{Die-Abenteuer-des-Huckleberry-Finn}. 
And since in every model there is only one, \emph{DL} sentences are satisfied. 
But this does not correspond to the intended constraint which is usual in \emph{CWA} way.   
%Classical constraint languages are based on the \emph{CWA} where constraints need to be satisfied only by named individuals which yields a constraint violation.
This ambiguity in semantics is one of the main reasons why \emph{OWL 2} has not been adopted as a standard constraint language for RDF validation in the past. 
In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (\emph{CWA} and \emph{UNA}) as for RDF validation in general. 
%In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (CWA) that is for RDF validation in general. 
%\tb{feedback from Erman: the major problem is semantics. OWL basically work in different semantics than others. And their semantics are formal. Closed world vs. Open World is also a part of this. / Thomas: I think your issue is tackled now? No.}

@@

The \emph{constraining element} indicates the actual type of constraint like \emph{DL} concept and role constructors, (in)equality, and further keywords for constraint types which cannot be expressed in \emph{DL} (e.g. regular expressions and constraining facets).
In some cases, a constraint is only complete when in addition to the \emph{constraining element} a \emph{constraining value} is stated.
The cardinality restriction 
\ms{$\geq$1 author.Person}, e.g., 
constructs an anonymous class of all individuals having at least one \emph{author} relationship to \emph{persons}.
The \emph{constraining element} of this \emph{property constraint} is the \emph{DL} at-least restriction \emph{$\geq$} and the \emph{constraining value} is \emph{1}.
\emph{Simple constraints} may refer to a list of \emph{classes}.
The qualified cardinality restriction above, e.g., refers to the class \emph{Person}, 
i.e. restricts the objects of the property \emph{author} to be of the class \emph{Person}.

For \emph{property constraints}, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths} (\emph{R-55}; also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} (\ms{authorOf(Stephen-Hawking, A-Brief-History-Of-Time)}) whose \emph{genre} is \emph{Popular-Science} ({\ms{genre(A-Brief-History-Of-Time, Popular-Science)}),
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} (\ms{authorOfGenre(Stephen-Hawking, Popular-Science)}):
\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}
Thus, when mapped to the \emph{RDF-CV} (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

As the \emph{The-Great-Gatsby} is a \emph{publication} (\ms{rdf:type(The-Great-Gatsby, Publication)}) having an \emph{ISBN} number (\ms{isbn(The-Great-Gatsby, '978-0241965672')}) and a \emph{title} (\ms{title(The-Great-Gatsby, 'The Great Gatsby')}) and not an additional \emph{ISSN} identifier, the \emph{The-Great-Gatsby} is a valid \emph{publication}.
%\begin{DL}
%rdf:type(BElannaTorres,Half-Klingon) \\
%klingonMother(BElannaTorres,Miral) \\
%klingonMother(BElannaTorres,JohnTorres)
%\end{DL}
The \emph{complex constraint} is mapped to the \emph{RDF-CV} (see Table \ref{tab:complex-constraints}) and expressed in \emph{DL} as follows:

@@

The \emph{DL} statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{(simple) generic constraint} conforming to the \emph{RDF-CV} (see Table \ref{tab:existential-quantification-as-simple-constraints}).
Then, one mapping (\emph{$m_1(sc_{OWL 2})$}, \emph{$m_2(sc_{ReSh})$}, \emph{$m_3(sc_{ShEx})$}, ...) is defined for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the corresponding \emph{generic constraint} (\emph{gc}).

>>>>

With \emph{minimum qualified cardinality restrictions} (\emph{R-75/81})\footnote{Requirements/constraint types are uniquely identified by alphanumeric technical identifiers like \emph{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}}, researchers from the library domain may restrict
that \emph{publications} must have at least one \emph{author} which must be a \emph{person}.
This constraint can be represented generically (\emph{generic constraint}) in \emph{DL} {\small\ms{(Publication $\sqsubseteq$ $\geq$1 author.Person)}}, but also specifically (\emph{specific constraint}) by multiple constraint languages:

@@

As every \emph{publication} must have at least one \emph{author} which must be a \emph{person}, 
and since the book \emph{The-Lord-Of-The-Rings} is a \emph{publication} {\small\ms{(rdf:type(The-Lord-Of-The-Rings, Publication)),}} 
it must have at least one \emph{author} relationship to a \emph{person}.
Constraint violations are raised either (1) if
\emph{The-Lord-Of-The-Rings} does not have any \emph{author} relationship, or (2) if 
it has an \emph{author} which is not a \emph{person} \\
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Hobbit)),}} 
or (3) if it has an \emph{author} for which no class is assigned
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)).}}
In contrast, \emph{The-Lord-Of-The-Rings} is a valid \emph{publication}, if it is connected to a \emph{person} via the property \emph{author}
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Person)).}}

@@

%The underlying \emph{semantics} for RDF validation is \emph{UNA} and \emph{CWA}.
RDF validation requires that different names represent different objects (\emph{unique name assumption (UNA)}), whereas,
\emph{OWL 2} is based on the {\em non-unique name assumption (nUNA)}.  
A data property is \emph{functional} (\emph{R-65}) if for each individual \emph{x}, there can be at most one distinct literal \emph{y} such that \emph{x} is connected by the data property to \emph{y}. 
When assuming \emph{UNA}, the functional property {\em title} {\small\ms{(funct (title))}} causes a clash
in case the book {\em Huckleberry-Finn} has more than one \emph{title}.
%(\ms{title(Huckleberry-Finn,} \ms{The-Adventures-of-Huckleberry-Finn)}, \ms{title} \ms{(Huckleberry-Finn,} \ms{Die-Abenteuer-des-Huckleberry-Finn)}).
Assuming \emph{nUNA}, however, reasoning concludes that the titles {\em The-Adventures-of-Huckleberry-Finn} and {\em Die-Abenteuer-des-Huckleberry-Finn} are the same. 
%(\ms{\{The-Adventures-of-Huckleberry-Finn\}} = \ms{\{Die-Abenteuer-des-Huckle} \ms{berry-}\ms{Finn\}}).

Reasoning in \emph{OWL 2} is based on the {\em open-world assumption (OWA)}, i.e., a statement cannot be inferred to be false if it cannot be proved to be true. 
%which fits its primary design purpose to represent knowledge on the World Wide Web. 
%Most of the constraint languages except OWL, have a major shortcoming of redundancy.
%With its open-world semantics, used as a constraint language, OWL solves the redundancy problem. 
%\tb{the following example does not show the redundancy problem!}
%\tb{2. benefit: reasoning solves redundancy problem}
As each \emph{book} must have a \emph{title} {\small\ms{(Book $\sqsubseteq \exists$ title.$\top$)}} and {\em Hamlet} is a \emph{book} {\small\ms{(Book(Hamlet)),}}
{\em Hamlet} must have at least one \emph{title}.
In an \emph{OWA} setting, this axiom does not cause a constraint violation (even if there is no explicitly defined \emph{title}), since there must be a \emph{title} for this \emph{book} which we may not know. 
On the other hand, RDF validation scenarios require the \emph{closed-world assumption (CWA)}, i.e., a statement is inferred to be false if it cannot be proved to be true.
When assuming the \emph{CWA}, a constraint violation is raised as there is no explicitly defined title. 
A constraint that says that a book must have at least one title in \emph{OWA}, 
does not mean that the title has to be unique. 
In every model  (an interpretation that makes a set of \emph{DL knowledge base} true), 
it can be the case that a book has a single title, however, 
this title is not unique. 
In some models it is \emph{The-Adventures-of-Huckleberry-Finn}, in others it is \emph{Die-Abenteuer-des-Huckleberry-Finn}. 
And since in every model there is only one, \emph{DL} sentences are satisfied. 
But this does not correspond to the intended constraint which is usual in \emph{CWA} way.   
%Classical constraint languages are based on the \emph{CWA} where constraints need to be satisfied only by named individuals which yields a constraint violation.
This ambiguity in semantics is one of the main reasons why \emph{OWL 2} has not been adopted as a standard constraint language for RDF validation in the past. 
In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (\emph{CWA} and \emph{UNA}) as for RDF validation in general. 
%In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (CWA) that is for RDF validation in general. 
%\tb{feedback from Erman: the major problem is semantics. OWL basically work in different semantics than others. And their semantics are formal. Closed world vs. Open World is also a part of this. / Thomas: I think your issue is tackled now? No.}

@@

The \emph{constraining element} indicates the actual type of constraint like \emph{DL} concept and role constructors, (in)equality, and further keywords for constraint types which cannot be expressed in \emph{DL} (e.g. regular expressions and constraining facets).
In some cases, a constraint is only complete when in addition to the \emph{constraining element} a \emph{constraining value} is stated.
The cardinality restriction 
{\small\ms{$\geq$1 author.Person}}, e.g., 
constructs an anonymous class of all individuals having at least one \emph{author} relationship to \emph{persons}.
The \emph{constraining element} of this \emph{property constraint} is the \emph{DL} at-least restriction \emph{$\geq$} and the \emph{constraining value} is \emph{1}.
\emph{Simple constraints} may refer to a list of \emph{classes}.
The qualified cardinality restriction above, e.g., refers to the class \emph{Person}, 
i.e. restricts the objects of the property \emph{author} to be of the class \emph{Person}.

For \emph{property constraints}, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the \emph{constraining element} - which may be an operator (e.g., $\sqsubseteq$ in case of \emph{object property paths}).
\emph{Object Property Paths} (\emph{R-55}; also called \emph{complex role inclusion axioms} in \emph{DL})
express that, if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the \emph{author} of the book \emph{A-Brief-History-Of-Time} {\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} whose \emph{genre} is \emph{Popular-Science} {\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the following \emph{object property path} infers that \emph{Stephen-Hawking} is an \emph{author} of the \emph{genre} \emph{Popular-Science} {\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science)):}}
\begin{DL}
authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre 
\end{DL}
Thus, when mapped to the \emph{RDF-CV} (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are on the left side of the \emph{constraining element} \emph{$\sqsubseteq$},
and the property \emph{authorOfGenre} is on the right side:

@@

As the \emph{The-Great-Gatsby} is a \emph{publication} {\small\ms{(rdf:type(The-Great-Gatsby, Publication))}} having an \emph{ISBN} number {\small\ms{(isbn(The-Great-Gatsby, '978-0241965672'))}} and a \emph{title} {\small\ms{(title(The-Great-Gatsby, 'The Great Gatsby'))}} and not an additional \emph{ISSN} identifier, the \emph{The-Great-Gatsby} is a valid \emph{publication}.
%\begin{DL}
%rdf:type(BElannaTorres,Half-Klingon) \\
%klingonMother(BElannaTorres,Miral) \\
%klingonMother(BElannaTorres,JohnTorres)
%\end{DL}
The \emph{complex constraint} is mapped to the \emph{RDF-CV} (see Table \ref{tab:complex-constraints}) and expressed in \emph{DL} as follows:

@@

The \emph{DL} statement {\small\ms{Publication $\equiv$ $\exists$ publisher.Publisher}} is mapped to a \emph{(simple) generic constraint} conforming to the \emph{RDF-CV} (see Table \ref{tab:existential-quantification-as-simple-constraints}).
Then, one mapping (\emph{$m_1(sc_{OWL 2})$}, \emph{$m_2(sc_{ReSh})$}, \emph{$m_3(sc_{ShEx})$}, ...) is defined for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the corresponding \emph{generic constraint} (\emph{gc}).
